# Code-a-thon Project Template

This project is intended to help you apply your skills in R in a way that would be conducive for collaborating or creating reproducible analyses.


## Data Used

Identify a dataset to analyze. Remember that if you share the data you must **not violate any privacy restrictions**

If you don't have your own data, options for places to find data are listed on https://daseh.org.

You might start by setting up an R Project, importing your data, previewing the data, etc. 

```{r}
library(tidyverse)
library(broom)
library(naniar)
library(maps)
library(usmap)
library(tidycensus)
dat <- read_csv("https://drive.usercontent.google.com/download?id=1V_1-6dfl88E6R_F3NbedqUCjmFabyMJ7&export=download&authuser=0&confirm=t&uuid=9b3505cb-8033-4650-be3e-24d2e914609c&at=AAINaIJgGDkjG47-KSbtWO1UTWTq:1782243211418")
dat
```

**describe** what your data looks like and **where** you got it. Identify what the variables and samples are, or point to a data dictionary. Describe how the data was originally created. 

US census data for Long Beach, California. Wide data with 26 columns. I got this cleaned US census data for Long Beach from Data Day 2026 at CSULB. Samples are surveyed individuals and the variables are; 
SERIALNO
Pwgtp 
Year
age
Sex
Location
Race_eth
Age_cat
Education
Personal_income
Household_income
Class
Has_medicaid
Has_health_insurance
Has_disability
Has_food_stamps
Tenure
Income_to_poverty_ratio
Below_poverty_line
Household
Hisp_Ethnicity
AAPI_Ethnicity

```{r}
nrow(dat)
ncol(dat)
colnames(dat)
miss_var_which(dat)
```


# Data Cleaning/Wrangling

Perform any data subsetting, cleaning, or manipulation methods that were described in this course on your data. Examples are: renaming the columns, recoding values, reshaping the data, filtering the data etc. 

You might also research other methods to include here, especially if your data is very unique.

```{r}
#data subset for 2023, exclude 2021
dat |> pull(year) |> range()
dat |> filter(year == 2021) |> nrow()
dat |> filter(year == 2022) |> nrow()
dat |> filter(year == 2023) |> nrow()
subdat <- dat |> filter(year == 2023)

subdat |> pull("has_medicaid") |> class()
#turn character columns into numeric to calculate percentages later
subdat <- subdat |> mutate(has_medicaid = if_else(has_medicaid == "Yes", 1L, 0L),
has_health_insurance = if_else(has_health_insurance == "Yes", 1L, 0L),
has_disability = if_else(has_disability == "Yes", 1L, 0L),
has_food_stamps = if_else(has_food_stamps == "Yes", 1L, 0L)
)

#public assistance column has lots of 0 and -1 which will mess with averages later, I'll replace with na and omit during calculations 
#In U.S. Census Bureau microdata a -1 is a missing data flag or a placeholder.
subdat <- subdat |> replace_with_na(replace = list(public_assistance_income = c(-1,0)))
subdat
```


# Data Visualization

Create some visualizations of your data using the esquisse app or the ggplot2 package.

```{r}
int_Country <- Happiness |> filter(Country_name %in% c("China", "United States", "India", "Mexico"))

ggplot(
  data = int_Country,
  aes(
    x = Country_name,
    y = Happiness_score
  ),
  group = Country_name
) +
  geom_point()

```


# Data Analysis

Perform a **simple analysis** of your data. This can involve summarizing the data to describe aspects about it (quartiles, means, range etc.) or a simple statistical test. 

Don't forget to describe what analysis you performed and why. Provide some simple **interpretation** about what your analysis might indicate about your data.

```{r}
dat |> pull(location) |> unique() 
#total amount surveyed individuals per lb neighborhood
east_total <- subdat |> filter(location == "East Long Beach") |> nrow()
south_total <- subdat |> filter(location == "South Long Beach") |> nrow()
north_total <- subdat |> filter(location == "North Long Beach") |> nrow()
central_total <- subdat |> filter(location == "Central Long Beach") |> nrow()

#calculate percentage of surveyed individuals on public assistance (medicaid, disability, food stamps) or have health insurance

#health insurance
health_e <- subdat |> filter(location == "East Long Beach") |> pull(has_health_insurance) |> sum() / east_total 
health_s <- subdat |> filter(location == "South Long Beach") |> pull(has_health_insurance) |> sum() / south_total
health_n <- subdat |> filter(location == "North Long Beach") |> pull(has_health_insurance) |> sum() / north_total
health_c <- subdat |> filter(location == "Central Long Beach") |> pull(has_health_insurance) |> sum() / central_total

percentage <- function(x) x * 100
health_e <- percentage(x = health_e)
health_s <- percentage(x = health_s)
health_n <- percentage(x = health_n)
health_c <- percentage(x = health_c)

#medicaid
med_e <- subdat |> filter(location == "East Long Beach") |> pull(has_medicaid) |> sum() / east_total 
med_s <- subdat |> filter(location == "South Long Beach") |> pull(has_medicaid) |> sum() / south_total
med_n <- subdat |> filter(location == "North Long Beach") |> pull(has_medicaid) |> sum() / north_total
med_c <- subdat |> filter(location == "Central Long Beach") |> pull(has_medicaid) |> sum() / central_total

med_e <- percentage(x = med_e)
med_s <- percentage(x = med_s)
med_n <- percentage(x = med_n)
med_c <- percentage(x = med_c)

#disability
able_e <- subdat |> filter(location == "East Long Beach") |> pull(has_disability) |> sum() / east_total 
able_s <- subdat |> filter(location == "South Long Beach") |> pull(has_disability) |> sum() / south_total
able_n <- subdat |> filter(location == "North Long Beach") |> pull(has_disability) |> sum() / north_total
able_c <- subdat |> filter(location == "Central Long Beach") |> pull(has_disability) |> sum() / central_total

able_e <- percentage(x = able_e)
able_s <- percentage(x = able_s)
able_n <- percentage(x = able_n)
able_c <- percentage(x = able_c)

#food stamps
food_e <- subdat |> filter(location == "East Long Beach") |> pull(has_food_stamps) |> sum() / east_total 
food_s <- subdat |> filter(location == "South Long Beach") |> pull(has_food_stamps) |> sum() / south_total
food_n <- subdat |> filter(location == "North Long Beach") |> pull(has_food_stamps) |> sum() / north_total
food_c <- subdat |> filter(location == "Central Long Beach") |> pull(has_food_stamps) |> sum() / central_total

food_e <- percentage(x = food_e)
food_s <- percentage(x = food_s)
food_n <- percentage(x = food_n)
food_c <- percentage(x = food_c)

#average personal income per neighborhood
money_e <- subdat |> filter(location == "East Long Beach") |> pull(personal_income) |> mean()
money_s <- subdat |> filter(location == "South Long Beach") |> pull(personal_income) |> mean()
money_n <- subdat |> filter(location == "North Long Beach") |> pull(personal_income) |> mean()
money_c <- subdat |> filter(location == "Central Long Beach") |> pull(personal_income) |> mean()

#average household income per neighborhood
house_e <- subdat |> filter(location == "East Long Beach") |> pull(household_income) |> mean()
house_s <- subdat |> filter(location == "South Long Beach") |> pull(household_income) |> mean()
house_n <- subdat |> filter(location == "North Long Beach") |> pull(household_income) |> mean()
house_c <- subdat |> filter(location == "Central Long Beach") |> pull(household_income) |> mean()


#how many people have public assistance income in long beach total: 80
public <- subdat |>  na.omit() |> nrow()

#percentage people have public assistance income in long beach per neighborhood
avg_pub_e <- subdat |> filter(location == "East Long Beach") |> na.omit() |> nrow() / east_total
avg_pub_e <- percentage(x = avg_pub_e)

avg_pub_s <- subdat |> filter(location == "South Long Beach") |> na.omit() |> nrow() / south_total
avg_pub_s <- percentage(x = avg_pub_s)

avg_pub_n <- subdat |> filter(location == "North Long Beach") |> na.omit() |> nrow() / north_total
avg_pub_n <- percentage(x = avg_pub_n)

avg_pub_c <- subdat |> filter(location == "Central Long Beach") |> na.omit() |> nrow() / central_total
avg_pub_c <- percentage(x = avg_pub_c)

#average public assistance income per neighborhood
public_e <- subdat |> filter(location == "East Long Beach") |> na.omit() |> pull(public_assistance_income) |> mean()
public_s <- subdat |> filter(location == "South Long Beach") |> na.omit() |> pull(public_assistance_income) |> mean()
public_n <- subdat |> filter(location == "North Long Beach") |> na.omit() |> pull(public_assistance_income) |> mean()
public_c <- subdat |> filter(location == "Central Long Beach") |> na.omit() |> pull(public_assistance_income) |> mean()
```


# Please include additional sections as desired.

```{r}

```


# Versions

Include some information about the version of R and the packages you are using by running sessionInfo().

```{r}
sessionInfo()
```

Render your document to an html or pdf using the "Knit" button.
